|

### 2020-05-19 paper/Manuscript.Rmd
some text on WORCS and DORA and FAIR

>>>>

WORCS encourages and simplifies the adoption of Open Science principles in daily scientific work. It helps to make all research objects that are created through the scientific process (data and code, but also manuscripts) open and publicly assessible. This enables other researchers to reproduce results, but also provides a proper basis for research evaluation according to the San Francisco Declaration on Research Assessment (DORA) [https://sfdora.org/], which is increasingly adopted by research institutions around the globe. For example, reviewers of grant proposals have immediate access to previous work of a researcher or a group of researchers, and can assess research quality based on content rather than relying on spurious surrogate indicators like personal h-indexes, journal impact factors and conference rankings. The detailed version control and commit tracking of github and similar platforms furthermore make it possible to assess the contributions made by different researchers, which can for example play a role in hiring or promotion procedures.

The FAIR Guiding Principles for scientific data management and stewardship [https://doi.org/10.1038/sdata.2016.18] advocate that all digital research objects should be Findable, Accessible, Interoperable and Reusable. They are increasingly adopted by research organizations and funders as a requirement towards more open, transparent and reproducible science. The WORCS workflow includes a step to publish the original raw data (if possible) or alternatively a synthetic data set in the (public) project repository, thus making the research data or a workable substitute accessible. WORCS uses CSV as an open and widely used, tabular format for data representation, thus enabling interoperability with other data sets and software. The R scripts for analyzing the data are naturally part of the same repository, facilitating reproducibility of obtained results by other researchers.

Next to data, also research software should be FAIR [https://doi.org/10.3233/DS-190026]. WORCS itself sets a good example and follows the recommendations for FAIR research software [https://fair-software.eu/]: It is hosted on a public repository with version control (github), it has a licence (GPL v3.0), it is registered in a community registry (CRAN), it enables the citation of the software (<!-- this paper. anything else? consider adding a cff file -->), and a software quality checklist was used for its development (<!-- which one? -->). By connecting the created R projects to github by default, it also guides its users through the first steps towards making their software FAIR, and lowers the threshold for taking the remaining steps.
